---
title: "Exploratory Data Analysis"

---

<h2>  Time Series Analysis   </h2>


Let's start with reading in the intraday range data. First with the S&P 500.

``` {r}
library(tidyverse)
library(quantmod)


spyIn <- quantmod::getSymbols("SPY", from = as.Date("2021/01/01"), to = as.Date("2023/09/30"), periodicity = "daily", src = "yahoo", auto.assign = FALSE)
qqqIn <- quantmod::getSymbols("QQQ", from = as.Date("2021/01/01"), to = as.Date("2023/09/30"), periodicity = "daily", src = "yahoo", auto.assign = FALSE)
iwmIn <- quantmod::getSymbols("IWM", from = as.Date("2021/01/01"), to = as.Date("2023/09/30"), periodicity = "daily", src = "yahoo", auto.assign = FALSE)


spyIn$spyRange <- (spyIn$SPY.High - spyIn$SPY.Low)/ spyIn$SPY.Open
qqqIn$qqqRange <- (qqqIn$QQQ.High - qqqIn$QQQ.Low)/ qqqIn$QQQ.Open
iwmIn$iwmRange <- (iwmIn$IWM.High - iwmIn$IWM.Low)/ iwmIn$IWM.Open


#decomposedSPY <- decompose(spyIn$spyRange)

##decompedSPY = HoltWinters(spyIn$spyRange,beta = FALSE,gamma = FALSE)
#plot(decompedSPY)

autoplot(spyIn$spyRange)
acf(spyIn$spyRange)
pacf(spyIn$spyRange)

```


Upon initial review of the SPY intraday range data, it appears that there is some trend, but no seasonality. The decomposition function would not work on the data as it could not recognize periodicity. As a next step, we can difference the data.

``` {r}
library(forecast)


diff1 <- diff(spyIn$spyRange)



#write.csv(diff1SPY, "/data/spyDiff.csv")
#auto.arima(spyIn$spyRange)
plot(diff1)
```

After differencing, the data appears to have lost its trend. However, there is still visible heteroskedacticity in the differences between intraday ranges.


``` {r}
acf(diff1,    na.action = na.exclude)

pacf(diff1,    na.action = na.exclude)

```

After differencing, we see 2 lags being significant in the ACF plot. In the PACF plot, we see about 4 lags being significant. This is a marked departure from the original plots, where the ACF showed clear non-stationarity and the PACF had many significant lags.

``` {r}
library(tseries)

diffnoNA <- diff1$spyRange[!is.na(diff1$spyRange)]
adf.test(diffnoNA)

#
```


With dickey fuller test result of 0.01, we can reject the null hypothesis and conclude the series is stationary. 



<h3> Moving Average Smoothing </h3>

```{r}

spyDiff1 <- as.ts(diffnoNA)

spyAvg3Diff1 <- stats::filter(spyDiff1, sides = 1, rep(1/3, 3))
spyAvg5Diff1 <- stats::filter(spyDiff1, sides = 1, rep(1/5, 5))
spyAvg30Diff1 <- stats::filter(spyDiff1, sides = 1, rep(1/30, 30))
spyAvg50Diff1 <- stats::filter(spyDiff1, sides = 1, rep(1/50, 50))

spyAvg100Diff1 <- stats::filter(spyDiff1, sides = 1, rep(1/100, 100))


autoplot(spyAvg3Diff1)

autoplot(spyAvg5Diff1)


autoplot(spyAvg30Diff1)


autoplot(spyAvg50Diff1)

autoplot(spyAvg100Diff1)

```


Overall the effect of the moving average at higher numbers is to reveal seasonality in the data. At lower smoothing levels (3 and 5 days), the time series looked almost unchanged, and still highly variant. Only at a 30 day average window did the seasonality start to appear, with clear periods in the data and a repeating pattern. Notably, even with a repeating pattern the data still showed obvious heteroskedasticity, with periods of increased varianced. The 50 and 100 day moving average windows began to obfuscate the periods, while keeping the heteroskedacticity, although they did have smaller variations and were nearer to 0 on average. In adition, the 50 and 100 days had discernable trends in the data.


